---
title: "Investigate: Competitiveness and seating preference"
output: github_document
---

Investigators: Ayla Pearson, Bonnie Zhang, George Wu, Socorro Dominguez

Date: April 6th, 2019

The current analysis seeks to explore the relationship between `self-reported competitiveness` and `seating preference` in a classroom. We also want to determine whether this relationship is confounded by an individual's `gender` or `enagement` level. Toward this goal, we constructed a [survey](https://ubc.ca1.qualtrics.com/jfe/form/SV_d1ex7dheBUxPGPb), and collected responses from individuals affiliated with the Master of Data Science program at the University of British Columbia.

## Data ingestion

We compiled the seating preference, self-reported competitiveness, gender, as well as engagement information from 64 survey entries. A full description of the dataset and ingestion procedures can be found [here](https://github.ubc.ca/sedv8808/seat_pref_survey).

```{r load libraries}
# load libraries
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gridExtra))
library(apaTables)
```

```{r load data, message = FALSE}
# load data
df <- read_csv("../seat_tidy.csv")

# convert categorical variables to factors
df <- df %>% 
  mutate_at(c("seat_pref", "gender"), factor)

# preview data
df %>% head()
```

Note that the column `seat-pref` corresponds to seating preference, `comp` corresponds to self-reported competitiveness, `gender` corresponds to gender, and `engage` corresponds to an individual's engagement level.

## Data overview

#### Missing data

We offered the option for participants not to respond to certain survey questions. We will simply consider these omitted responses as missing values.  

```{r evaluate missing data}
# re-encode "prefer not to response" (level 4) as missing data
df$seat_pref <- df$seat_pref %>% 
  fct_recode("Front" = "1", 
             "Middle" = "2", 
             "Back" = "3",  
             NULL = "4")

df$gender <- df$gender %>% 
  fct_recode("Man" = "1",
             "Woman" = "2",
             NULL = "4")

# evaluate missing data
df %>% summary()
```

We are missing 4 entries for `seating preference`, 5 entries for `gender`, as well as 1 entry for `engagement`. As we do not have enough information to meaningfully impute these missing values, we will simply drop these missing records from the dataset.

Note also that there were no participants that identified with genders other than man or woman.

```{r drop missing data}
# drop missing data
df <- df %>% drop_na()
df %>% dim()
```

We have 56 remaining records after removing the missing information.

#### Variable scaling

Both continuous variables, `comp` and `engage`, were measured using a 5-point Likert scale, as such, there should be no scaling issues for these two variables.

#### Multicollinearity

We want to make sure that the two continuous variables, namely the competitiveness variable (`comp`) and the engagement variable (`engage`), are not measuring the same psychological construct. For this reason, we will examine the Pearson correlation coefficient between these two variables.

```{r evaluate multicollinearity}
# evaluate correlation between comp and engage
# note that for EDA, we are not calculating the p-value for this coefficient
cor(x = df$comp, y = df$engage, method = "pearson")
```

It appears that self-reported competitiveness is rather weakly and positively associated with engagement level. This association is weak enough that we needn't worry too much about multicollinearity issues for this dataset. They seem to be indeed measuring different constructs.

## Univariate analysis

Let's take a closer look at each variable individually. Specifically, let's visualize the distribution of each variable.

#### Visual univariate analysis

```{r visualize univariate distribution}
# visualize seating preference
g31 <- df %>%   
  ggplot(aes(x = seat_pref)) +
  geom_bar() +
  theme_minimal() +
  xlab("Seating preference") +
  ylab("Number of people")

# visualize competitiveness
g32 <- df %>% 
  ggplot(aes(x = comp)) +
  geom_histogram(bins = 5, binwidth = 1) +
  theme_minimal() +
  xlab("Self-reported competitiveness") +
  ylab("Number of people")

# visualize gender
g33 <- df %>%   
  ggplot(aes(x = gender)) +
  geom_bar() +
  theme_minimal() +
  xlab("Gender") +
  ylab("Number of people")

# visualize engagement
g34 <- df %>% 
  ggplot(aes(x = engage)) +
  geom_histogram(binwidth = 1) +
  theme_minimal() +
  xlab("Engagement") +
  ylab("Number of people")

# display distribution of each variable
grid.arrange(g31, g32, g33, g34)
```

#### Numerical univariate analysis

We can also examine each variable numerically.

```{r descriptive univariate statistics}
# obtain descriptive statistics of each variable
df %>% summary()
```

Both visual and numeric examinations of each variable has led to the following observations:

Many participants prefer to sit in the middle of the classroom (n = 26), and fewer participants prefer sitting in the front (n = 18), or in the back (n = 12). Note that the number of participants in each response level is somewhat unbalanced.

The distribution of self-reported competitiveness is largely normal with a mean of 3. There are very few (less than five) participants with a score of 1 or with a score of 5.

There are slightly more self-identified men (n = 33) than self-identified women (n = 23) being included in the current analysis. The number of each gender is very roughly balanced (not terribly unbalanced).

The distribution of engagement level is slightly left-skewed with a mean of about 3.7. There are no participants associated with a score below 1. We believe the skewness of this distribution does not warrant major data transformations.

#### Transformation: seating preference as a binary variable

To address the unbalanced nature of the `seat_pref` variable, it might be helpful to further consolidate the levels of seating preference. Specifically, we will bin the levels of `Front` and `Back` to a new level called `Not-middle`. This new binary seating preference variable will be used along side our original `seat_pref` for additional comparisons. 

```{r transform seat_pref to binary}
# bin front/back to a new level "not-middle"
df$seat_pref_binary <- df$seat_pref %>% 
  fct_collapse("Not-middle" = c("Front", "Back"))

df %>% 
  group_by(seat_pref_binary) %>% 
  summarise(n = n())
```

This new binary seating preference variable is now more balanced, with more than 26 participants in each level.

## Multivariate analysis

Let's turn our attention to the relationship among the variables. Recall that we are most interested in the relationship between seating preference (`seat_pref`) and self-reported competitiveness (`comp`).

#### Seating preference and competitiveness

Let's visualize the relationship between these two variables.

```{r visualize seat_pref and comp}
# visualize seat_pref and competitiveness
g41 <- df %>% 
  ggplot(aes(x = seat_pref, y = comp)) +
  geom_boxplot() +
  coord_flip() +
  xlab("Seating preference") +
  ylab("Competitiveness") +
  ggtitle("Seating preference and self-reported competitiveness") +
  theme_minimal()

# visualize binary seat_pref and competitiveness
g42 <- df %>% 
  ggplot(aes(x = seat_pref_binary, y = comp)) +
  geom_boxplot() +
  coord_flip() +
  xlab("Seating preference") +
  ylab("Competitiveness") +
  ggtitle("Seating preference (binary) and self-reported competitiveness") +
  theme_minimal()

# display plots
grid.arrange(g41, g42, ncol = 1)
```

Let's also examine the relationship between these two variables numerically.

```{r numerical analysis for seat_pref and comp}
# numeric analysis for original seat_pref
df %>% 
  group_by(seat_pref) %>% 
  summarise(n = n(),
            mean_competitiveness = mean(comp),
            sd_competitiveness = sd(comp))

# numeric analysis for binary seat_pref
df %>% 
  group_by(seat_pref_binary) %>% 
  summarise(n = n(),
            mean_competitiveness = mean(comp),
            sd_competitiveness = sd(comp))
```

Both visual and numeric examinations of these two variable has led to the following observations:

There are a wide variation of self-reported competitiveness within each category of seating preference, as suggested by the relatively high standard deviation (sd) of competitiveness for each category.

A faint pattern seems to be: participants who reported to be more competitive tend to prefer to sit in either the front or back of the classroom, whereas participants who reported to be less competitive seem to prefer to sit in the middle of the classroom. Using the new binary `seat_pref` can largely reflect this pattern as well.

To simplify downstream analysis, we will replace the original three-level `seat_pref` variable with our new binary `seat_pref_binary` variable.

#### Seating preference and competitiveness with gender

```{r visualize seat_pref and comp by gender}
# visualize seat_pref and competitiveness with gender 
df %>% 
  ggplot(aes(y = comp, x = seat_pref_binary)) +
  geom_boxplot() +
  coord_flip() +
  facet_wrap(~gender, ncol = 1) +
  xlab("Seating preference") +
  ylab("Competitiveness") +
  ggtitle("Seating preference and self-reported competitiveness by gender") +
  theme_minimal()
```

```{r numerical analysis for visualize seat_pref and comp by gender}
df %>% 
  group_by(seat_pref_binary, gender) %>% 
  summarise(n = n(),
            mean_competitiveness = mean(comp),
            sd_competitiveness = sd(comp))
```

Both visual and numerical examinations suggests that gender does not seem to greatly influence the relationship between competitiveness and seating preference. People who rated themselves to be more competitive tend to prefer not to sit in the middle of the class, for both genders.

#### Seating preference and competitiveness with engagement

To facilitate and further simplify comparisons, we will categorize participants with either high or low engagement using the median engagement score.

```{r categorize engagement}
# sort median engagement
median_engage = median(df$engage)

# categorize  engagement
df <- df %>% 
  mutate(engage_binary = if_else(engage > median_engage, "Highly-engaged", "Lowly-engaged"))

# convert to factor
df$engage_binary <- factor(df$engage_binary)

df %>%
  group_by(engage_binary) %>% 
  summarise(n = n())
```

We can then visualize the engagement along with seating preference and competitiveness. 

```{r}
# visualize seat_pref and competitiveness with gender 
df %>% 
  ggplot(aes(y = comp, x = seat_pref_binary)) +
  geom_boxplot() +
  coord_flip() +
  facet_wrap(~engage_binary, ncol = 1) +
  xlab("Seating preference") +
  ylab("Competitiveness") +
  ggtitle("Seating preference and self-reported competitiveness by engagement") +
  theme_minimal()
```

```{r numerical analysis for visualize seat_pref and comp by engagement}
df %>% 
  group_by(seat_pref_binary, engage_binary) %>% 
  summarise(n = n(),
            mean_competitiveness = mean(comp),
            sd_competitiveness = sd(comp))
```

Both visual and numerical examinations suggests that engagement does somewhat influence the relationship between competitiveness and seating preference. People who rated themselves to be more competitive tend to prefer not to sit in the middle of the class, especially they are also highly-engaged. Interestingly, some competitive participants who are not typically engaged in the class content, prefer sitting in the middle of class instead.

## Summary

The current analysis aims to explore the relationship between self-reported competitiveness and seating preference. We examined the distribution of each variable in the dataset, as well as the relationship and potential interactions among these variables, and transformed the variables as necessary. 

Using both visual and numeric examinations, we mainly observed that:

- Participants who rated themselves to more competitive tend to prefer sitting in either the front or back of the classroom, whereas participants who rated themselves to less competitive tend to prefer sitting in the middle of the classroom.

- Gender does not seem to influence the relationship between self-reported competitiveness and seating preference.

- Some participants who rated themselves to be more competitive prefer sitting in the middle of the class instead, if they are not engaged in the course content.

The current investigations suffer from a number of limitations, specifically:

- Participants are allowed to choose multiple seating preference, this means a small number of entries actually contain duplicated information. In other words, we actually have less information that we believe.

- The observed effect size of competitiveness across each level of seating preference is quite small. Furthermore, the variation of competitiveness is quite large within each category of seating preference.

- To simplify comparisons, we somewhat arbitrarily binarized the seating preference variable, as well as the engagement variable. Doing so can diminish the already weak effect size. 

This investigation is largely exploratory in nature, and will require further more robust statistical analyzes downstream.