---
title: "EDA_Seat"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

```{r}
data = read_csv('data/seat_tidy.csv')

data$gender <- as.factor(data$gender)
data$seat_pref <- as.factor(data$seat_pref)
data$gender = fct_recode(data$gender, Woman = "1", Man = "2", Not_answer = "4")
data$seat_pref = fct_recode(data$seat_pref, Front = "1", Middle = "2", Back = "3", Not_answer = "4")

```

##### Bar graph of the count of different seat preferences
```{r}
data %>%
  ggplot(aes(factor(seat_pref), fill = seat_pref)) +
  geom_bar() +
  theme() +
  xlab("Seating Preference")
```

**Interpretation**: We can see there are about 20 respondents choose to sit in the front, 28 respondents choose to sit in the middle and 12 respondents choose to sit in the back. Individuals who preferred to not disclosed their `Seat Preference` are 4.

#### Bar graph of the count of different seat preferences by gender
```{r}
data %>%
  ggplot(aes(factor(seat_pref),fill = gender)) +
  geom_bar() +
  theme() +
  xlab("Seating Preference")
```

**Interpretation:** The count of woman who prefer to sit in the front is similar that who prefer to sit in the middle. Man have stronger preference to sit in the middle comparing to the front and back seats. Majority of man choose to sit in the middle.

#### box plot of engagement by different seat preference  
```{r}
data %>%
  filter(!is.na(engage)) %>%
  ggplot(aes(seat_pref,engage, fill = seat_pref)) +
  geom_boxplot() +
  theme() +
  xlab("Seating Preference")
```

**Interpretation:** Individuals who prefer to sit in the front has the highest average engagement level around 3.8, while individuals who prefer to sit in the middle and back has similar average engagement level around 3.7. The engagement average is pretty similar same for the three seat preferences.We also notice that the average engagement level of individuals who choose to sit in the front is more spread out comparing to the other two seat preferences. For individuals who choose not to answer, we do not really have a plot that throws much information.

#### box plot of competitiveness by different seat preference 
```{r}
data %>%
  ggplot(aes(seat_pref,comp, fill = seat_pref)) +
  geom_boxplot() +
  theme() +
  xlab("Seating Preference")
```

**Interpretation:** Individuals who prefer to sit in the back has the highest average competitivenes level around 3.8, while individuals who prefer to sit in the front and middle has similar average competitivenes level around 3.0. For individuals who choose not to answer, we do not really have a plot that throws much information.



