---
title: "EDA_Gender"
author: "Seating Preference Team"
date: '2019-04-05'
output: github_document
---

```{r setup, include = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
```
## Research Question of Interest

`Seating location` in the classroom is commonly linked to grade performance, student engagement, and learning experience (Gowan et al., 2017; Shernoff, et al., 2017). Nevertheless, the question remains whether certain personality dispositions systematically influence where students prefer to sit in a classroom. We are asking a group of MSD students in order to learn more about the relationship between `seating location` and competitiveness. 

## Gender as a Cofound

In our survey, we are also wanting to take into consideration confound variables and we think `Gender` is one, since `Gender` has been directly linked to both student seating (Shernoff, et al., 2017)

### Some Tables

```{r summary tables}
data <- read_csv("data/seat_tidy.csv")

data %>%
  group_by(gender) %>%
  na.omit() %>%
  summarize(count = n(), competitiveness = mean(comp), engagement = mean(engage))
```
**Interpretation:** We can see that individuals that identified themselves as `Women` have the highest `Competitiveness` average while individuals that prefer to not disclose their gender had the lowest. 

The previous table was merely used considering `Gender` without taking into consideration where each individual likes seating. Let's explore that now:

```{r summary tables 2}
data %>%
  group_by(gender, seat_pref) %>%
  summarize(count = n(), competitiveness = mean(comp), engagement = mean(engage))
```
**Interpretation:** Individuals gender seems to be somewhat linked to where they prefer to seat AND to their competitiveness.

Women seem to be the most "average" competitors in terms that, no matter where they seat, they share a high level of competitiveness, all above 3 points. Even women who did not disclose where they would sit, had a high score of 3. (Men who did not want to disclose their preferred seat had a score of 1.)

For Men, Individuals seating at the front had the lowest competitiveness average, same as individuals in the middle. However, self reported men seating at the back end had the highest competitiveness levels of the whole sample. Surprisingly, men who did not disclose where they preferred seating had the lowest competitiveness average, same as people who did not disclose gender but preferred seating in the middle.

Individuals who did not disclose their gender but preferred seating in the front had a high competitiveness level; very different from those who preferred seating in the middle (lowest competitiveness level).

I believe `Gender` is definitely linked to competitiveness and seating preference.

Engagement is another variable that we studied, and we can see that most `Men` identified inidividuals had a really high engagement rate. Even groups who had a low `Competitiveness` average, showcased a very high `Engagement` score.  Women's ranged from 3.5 to 4; Men's from 3 to 4, Not identified, from 2.8 to 4.6; the highest engagement value!! However, we have to be very careful when block-ing too much the values since we can see that for some groups of people, our count is just of 1 individual, not enough data.

## Including Plots

The most basic plot we can do from here, is to do a count of the `Gender` of our respondents:
```{r gender count}
data %>%
  ggplot(aes(factor(gender))) +
  geom_bar() +
  theme_bw() +
  labs(title = "Gender Count", 
       x = "Self-identified Gender",
       y = "Count")
```

**Interpretation:** We can see that we have about 35 `Women` respondents while just 25 `Men` respondents. Individuals who preferred to not disclosed their `Gender` are 5. 

It would be interesting to see how this count reflects in their seat preference:

```{r gender count by seat preference}
data %>%
  ggplot(aes(factor(gender))) +
  geom_bar() +
  facet_wrap(~seat_pref) +
  theme_bw() +
  labs(title = "Gender Count as per Seat Preference", 
       x = "Self-identified Gender",
       y = "Count")
```
**Interpretation:** Women seem more likely to seat either at the fron or middle. Men seem to prefer the middle. People who would not disclose their gender, would also prefer to not disclose their preferred seat. 


We would like now to explore what role does the `Gender` play in an individuals `Engagement`

```{r engagement and gender}
data %>%
  filter(!is.na(engage)) %>%
  ggplot(aes(gender, engage)) +
  geom_boxplot() +
  theme_bw() +
  labs(title = "Gender vs Engagement",
       x = "Gender",
       y = "Engagement")
```
**Interpretation:** Men's `Engagement` seems to have a smaller standard deviation, with most of them having an `Engagement` score between 3.3 and 4.1; Women rated their engagement anywhere from 3 to almost 4.4 while Unidentified individuals ranged from 2.4 to 3.6. Although `men` seem to have similar patterns of engagement, some females scored higher in this aspect.

Same as before, it is normal to wonder where are these people seating:


```{r competitiveness and gender}
data %>%
  ggplot(aes(gender, comp)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.2) +
  theme_bw() +
  labs(title = "Gender vs Competitiveness",
       x = "Gender",
       y = "Competitiveness")
```


**Interpretation:** It is interesting to see that the competitiveness average is pretty similar same for the three categories of `Gender` we have. However, women show a larger span in terms of How competitive they perceive themselves to be, while men had a shorter span. Men's mean percentile is also closer to their 75 percentile. For individuals who did not report their gender, we do not really have a plot that throws much information; we need more individuals.